Validate Structural Analysis \\ \large Testing GPV Method in First-Price Reverse Auctions

### 2020-12-08 paper/paper.tex
fix typos

The central assumption underlying structural estimation is that the model used 
is correct in a particular environment. To overcome such a drawback, alternative 
studies such as lab experiments are used to confirm the theory's validity. But 
laboratory environments ultimately differ from the real environments and 
the conclusions from laboratory studies may not hold in more realistic environments. 
The goal of this paper is to test model validity in a new different way by 
borrowing the techniques from statistical learning literature. In statistical 
learning, multiple models for predictive purposes are trained on part of the dataset 
and their performance is evaluated on another reserved split of 
the available dataset. The technique is called cross-validation 
and is used when there are candidate models and their performance needs to 
be compared to select the fittest one. In this paper, I will adapt the idea 
of cross-validation to test the predictiveness of economic models. By splitting 
data into training data and test data as in statistical learning, I will 
run structural estimation on the training data, simulate test data using the 
results from strucutral estimation, and compare the simulated data to actual 
test data. If the difference is statistically significant, that suggests 
unfitness of the model in the particular environment. 
In this paper, I will use data from government procurement auctions to test 
the first-price sealed-bid equilibrium model \cite{RileySamuelson1981} 
for structural estimation
\footnote{The original goal of this paper also includes comparing the cost 
distribution of DBE/MBE/WBE (standing respectively for Disadvantaged/Minority/Woman
Business Enterprise) to the cost distribution of other firms, which will add 
to the existing literature on understanding how DBE/MBE/WBE perform relative 
to other firms. But due to data availability on these types of firms 
and time constraints on the project as this is a 
term paper, this goal is omitted. }.

Similar work has been done in \citeauthor{ChernomazYushioto2019} 
\citeyear{ChernomazYushioto2019}, where they have compared the accuracy 
of semiparametric and nonparametric structural estimation methods in an 
asymetric environment.

@@

This two-step estimation method has the advantage that in order to estimate cost 
distribution with observed bids, I do not need to assume a prior distribution 
of $f(c)$ to estimate it (which is a parametric method that works by 
first assuming the shape of a distribution and then estimating parameters of the 
distribution from data).

@@

\section[title]{Simulation} 
To test if the estimated cost distribution captures the underlying cost 
distribution that induces observed bids, I will proceed to validation stage: 
simulate bids using estimated cost distribution 
$F_c(\cdot)$ and comparing simulated bids to actual bids. For the test data, 
I will use an auction from the dataset which has 16 bidders. The data is not used 
to estimate the cost distribution because the total number of bids with 
$16$ bidders is less than $32$.

@@

Figure \ref{fig:simulation} plos simulated costs against simulated bids that are 
inferred from simulated costs using equation (\ref{eqn:BiddingFunction}). 
The simulated costs and bids center around 110\% 
and the range is small, with the max and min differing by less than 4\%. This 
is due to the estimated unimodal cost distribution 
concentrating in the range [80, 200], with the peak around 110\%. 
With the inversion method repeated for 1000 times, it is natural that 
the costs and bids fall into the place where most bids and costs are 
concentrated. With the number of runs curtailed to 100 and 50, as shown in 
Figure \ref{fig:simulation_run_times}, the bids become more spread out and 
closer to the bid distribution. This shows that the number of runs is 
an important factor to determine in roder to have a more accurate simulation 
but it is outside the scope of this paper. For simplicity, I will keep 
referencing the data generated with 1000 runs.

@@

Figure \ref{fig:actualVsimulation} plots test data against simulated bids. 
A wide difference between simulation and test data is observed here. 
This difference, however, is not unexpected for two reasons: 
(1) the test data, which contains an auciton with 16 bidders, does not conform 
to the the bid distribution depicted in Figure \ref{fig:BidFrequency}, and 
(2) I am comparing most likely bids predicted by the estimated cost distribution 
to a particular auction. Therefore, the difference here not suggest the 
inaccuracy of the estimated cost distribution or the equilibrium model. 
Instead, this exercise suggests that a larger test dataset is needed 
to deliver more informative results.

\section{Conclusion}
This paper tries to propose a way to validate models used in structural analysis 
by cross-validation. The dataset is split into training data and test 
data.
First, this paper conducts structural estimation on a government 
procurement auction dataset about highway constructions and estimates the 
cost distribution on the training data using a two-step nonparametric procedure.
Second, using the 
estimated cost distribution, the paper then constructs simulated bids with the 
inversion method and compare the simulation to test data. 
But the wide difference between the simulation and test data does not suggests
that the first-price sealed-bid equilibrium model introduced in 
\citeauthor{RileySamuelson1981} \citeyear{RileySamuelson1981}
is not a suitable model to predict bidder's 
behavior in real-world auctions. Instead, to draw a firm concludesion requires 
more data.

>>>>

The central assumption underlying structural estimation is that the model used 
is correct in a particular environment. To overcome such a drawback, alternative 
studies such as lab experiments are used to confirm the theory's validity. But 
laboratory environments ultimately differ from the real environments and 
the conclusions from laboratory studies may not hold in more realistic environments. 
The goal of this paper is to test model validity in a new different way by 
borrowing the techniques from statistical learning literature. In statistical 
learning, multiple models for predictive purposes are trained on part of the dataset 
and their performance is evaluated on another reserved split of 
the available dataset. The technique is called cross-validation 
and is used when there are candidate models and their performance needs to 
be compared to select the fittest one. In this paper, I will adapt the idea 
of cross-validation to test the predictiveness of economic models. By splitting 
data into training data and test data as in statistical learning, I will 
run structural estimation on the training data, simulate test data using the 
results from structural estimation, and compare the simulated data to actual 
test data. If the difference is statistically significant, that suggests 
unfitness of the model in the particular environment. 
In this paper, I will use data from government procurement auctions to test 
the first-price sealed-bid equilibrium model \cite{RileySamuelson1981} 
for structural estimation
\footnote{The original goal of this paper also includes comparing the cost 
distribution of DBE/MBE/WBE (standing respectively for Disadvantaged/Minority/Woman
Business Enterprise) to the cost distribution of other firms, which will add 
to the existing literature on understanding how DBE/MBE/WBE perform relative 
to other firms. But due to data availability on these types of firms 
and time constraints on the project as this is a 
term paper, this goal is omitted. }.

Similar work has been done in \citeauthor{ChernomazYushioto2019} 
\citeyear{ChernomazYushioto2019}, where they have compared the accuracy 
of semiparametric and nonparametric structural estimation methods in an 
asymmetric environment.

@@

This two-step estimation method has the advantage that to estimate cost 
distribution with observed bids, I do not need to assume a prior distribution 
of $f(c)$ to estimate it (which is a parametric method that works by 
first assuming the shape of a distribution and then estimating parameters of the 
distribution from data).

@@

\section[title]{Simulation} 
To test if the estimated cost distribution captures the underlying cost 
distribution that induces observed bids, I will proceed to the validation stage: 
simulate bids using estimated cost distribution 
$F_c(\cdot)$ and comparing simulated bids to actual bids. For the test data, 
I will use an auction from the dataset which has 16 bidders. The data is not used 
to estimate the cost distribution because the total number of bids with 
$16$ bidders is less than $32$.

@@

Figure \ref{fig:simulation} plots simulated costs against simulated bids that are 
inferred from simulated costs using equation (\ref{eqn:BiddingFunction}). 
The simulated costs and bids center around 110\% 
and the range is small, with the max and min differing by less than 4\%. This 
is due to the estimated unimodal cost distribution 
concentrating in the range [80, 200], with a peak around 110\%. 
With the inversion method repeated 1000 times, it is natural that 
the costs and bids fall into the place where most bids and costs are 
concentrated. With the number of runs curtailed to 100 and 50, as shown in 
Figure \ref{fig:simulation_run_times}, the bids become more spread out and 
closer to the bid distribution. This shows that the number of runs is 
an important factor to determine to have a more accurate simulation 
but it is outside the scope of this paper. For simplicity, I will keep 
referencing the data generated with 1000 runs.

@@

Figure \ref{fig:actualVsimulation} plots test data against simulated bids. 
A wide difference between simulation and test data is observed here. 
This difference, however, is not unexpected for two reasons: 
(1) the test data, which contains an auction with 16 bidders, does not conform 
to the bid distribution depicted in Figure \ref{fig:BidFrequency}, and 
(2) I am comparing most likely bids predicted by the estimated cost distribution 
to a particular auction. Therefore, the difference here does not suggest the 
inaccuracy of the estimated cost distribution or the equilibrium model. 
Instead, this exercise suggests that a larger test dataset is needed 
to deliver more informative results.

\section{Conclusion}
This paper tries to propose a way to validate models used in structural analysis 
by cross-validation. The dataset is split into training data and test 
data.
First, this paper conducts structural estimation on a government 
procurement auction dataset about highway constructions and estimates the 
cost distribution on the training data using a two-step nonparametric procedure.
Second, using the 
estimated cost distribution, the paper then constructs simulated bids with the 
inversion method and compare the simulation to test data. 
But the wide difference between the simulation and test data does not suggest
that the first-price sealed-bid equilibrium model introduced in 
\citeauthor{RileySamuelson1981} \citeyear{RileySamuelson1981}
is not a suitable model to predict bidder's 
behavior in real-world auctions. Instead, to draw a firm conclusion requires 
more data.
